$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-07-17 paper/appendix.tex
Rename substitution to renaming and swapping to exchange

\begin{nilemma}
  \label{lm:subst-unused}
  For every variables $i$ and $j$, if $i \not\equiv j$ then $\Unused_i (\subst{P}{j}{i})$.
\end{nilemma}
\begin{proof}
  By structural induction on \textsc{Process} and \textsc{Var}.
\end{proof}

@@

\begin{nilemma}
  \label{lm:swap-swap}
  The function $\func{swap}_i \; P$ is its own inverse: $\func{swap}_i \; (\func{swap}_i \; P) \equiv P$.
\end{nilemma}
\begin{proof}
  By structural induction on \textsc{Process} and \textsc{Var}.
\end{proof}

@@

\begin{proof}[Proof (Sketch)]
  The proof is by induction on \textsc{Equals} $\eq{}$.
  Here we only consider the nontrivial base cases for $\constr{struct}$ and their symmetric variants.
  Full proof in \cite{Zalakain2020Agda}.
  We proceed by induction on \textsc{StructCong} $\eqeq$:
  \begin{itemize}
    \item
      Case $\constr{comp-assoc}$: trivial, as leftover typing is naturally associative.
    \item
      Case $\constr{comp-sym}$ for $\comp{P}{Q}$: we use framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of $Q$ to the one of $P$.
    \item
      Case $\constr{comp-end}$: trivial, as the typing rule for $\PO$ has the same input and output contexts.
    \item
      Case $\constr{scope-end}$: we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
      In the opposite direction, we instantiate the newly created channel to a type $\unit$ and a usage annotation $\lz$..
    \item
      Case $\constr{scope-ext}$ for $\new \comp{P}{Q}$: we to show that $P$ preserves the usage annotations of the unused variable (\autoref{lm:types-unused}) and then use strengthening (\autoref{thm:strengthening}).
      In the reverse direction, we use weakening (\autoref{thm:weakening}) on $P$ and show that lowering and then lifting $P$ results in $P$ (\autoref{lm:lower-lift}).
    \item
      Case $\constr{scope-comm}$:
      we use swapping (\autoref{thm:swapping}), and for the reverse direction swapping and \autoref{lm:swap-swap} to show that swapping two elements in $P$ twice leaves $P$ unchanged. \qedhere
  \end{itemize}
\end{proof}

\section{Substitution}
\label{app:subst-generalization}

\begin{nitheorem}[Generalised substitution]
  \label{thm:subst-generalization1}
  Let process $P$ be well-typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
  The substituted variable $i$ is capable of $m$ in $\Gamma_i$, and capable of $n$ in $\Psi_i$.
  Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
  In other words, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
  \begin{multicols}{2}
  \begin{itemize}
    \item $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$
    \item $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$
  \end{itemize}
  \end{multicols}
  Let $\Gamma$ and $\Psi$ be related such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
  Let $\Delta$ have a usage annotation $\lz$ at position $i$, so that all consumption from $m$ to $n$ must happen in $P$.
  Then substituting $i$ with $j$ in $P$ will be well-typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
  Refer to \autoref{fig:subst} for a diagramatic representation.
\end{nitheorem}

>>>>

\begin{nilemma}
  \label{lm:renaming-unused}
  For every variables $i$ and $j$, if $i \not\equiv j$ then $\Unused_i (\subst{P}{j}{i})$.
\end{nilemma}
\begin{proof}
  By structural induction on \textsc{Process} and \textsc{Var}.
\end{proof}

@@

\begin{nilemma}
  \label{lm:exchange-exchange}
  The function $\func{exchange}_i \; P$ is its own inverse: $\func{exchange}_i \; (\func{exchange}_i \; P) \equiv P$.
\end{nilemma}
\begin{proof}
  By structural induction on \textsc{Process} and \textsc{Var}.
\end{proof}

@@

\begin{proof}[Proof (Sketch)]
  The proof is by induction on \textsc{Equals} $\eq{}$.
  Here we only consider the nontrivial base cases for $\constr{struct}$ and their symmetric variants.
  Full proof in \cite{Zalakain2020Agda}.
  We proceed by induction on \textsc{StructCong} $\eqeq$:
  \begin{itemize}
    \item
      Case $\constr{comp-assoc}$: trivial, as leftover typing is naturally associative.
    \item
      Case $\constr{comp-sym}$ for $\comp{P}{Q}$: we use framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of $Q$ to the one of $P$.
    \item
      Case $\constr{comp-end}$: trivial, as the typing rule for $\PO$ has the same input and output contexts.
    \item
      Case $\constr{scope-end}$: we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
      In the opposite direction, we instantiate the newly created channel to a type $\unit$ and a usage annotation $\lz$..
    \item
      Case $\constr{scope-ext}$ for $\new \comp{P}{Q}$: we to show that $P$ preserves the usage annotations of the unused variable (\autoref{lm:types-unused}) and then use strengthening (\autoref{thm:strengthening}).
      In the reverse direction, we use weakening (\autoref{thm:weakening}) on $P$ and show that lowering and then lifting $P$ results in $P$ (\autoref{lm:lower-lift}).
    \item
      Case $\constr{scope-comm}$:
      we use exchange (\autoref{thm:exchange}), and for the reverse direction exchange and \autoref{lm:exchange-exchange} to show that exchanging two elements in $P$ twice leaves $P$ unchanged. \qedhere
  \end{itemize}
\end{proof}

\section{Renaming}
\label{app:renaming-generalization}

\begin{nitheorem}[Generalised renaming]
  \label{thm:subst-generalization1}
  Let process $P$ be well-typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
  The substituted variable $i$ is capable of $m$ in $\Gamma_i$, and capable of $n$ in $\Psi_i$.
  Renaming will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
  In other words, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
  \begin{multicols}{2}
  \begin{itemize}
    \item $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$
    \item $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$
  \end{itemize}
  \end{multicols}
  Let $\Gamma$ and $\Psi$ be related such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
  Let $\Delta$ have a usage annotation $\lz$ at position $i$, so that all consumption from $m$ to $n$ must happen in $P$.
  Then renaming $i$ to $j$ in $P$ will be well-typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
  Refer to \autoref{fig:subst} for a diagramatic representation.
\end{nitheorem}

### 2020-07-17 paper/paper.tex
Rename substitution to renaming and swapping to exchange

  We present the syntax, semantics, type system and corresponding type safety properties.
  For the first time in the \picalc{}, we use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates linearity constraints into process continuations.
  We generalise the algebras on multiplicities, allowing the user to choose a mix of linear, graded and shared typing.
  We provide framing, weakening and strengthening theorems, which can now be stated for the linear \picalc{} and use them to prove subject congruence.
  We show that the type system is stable under substitution and prove subject reduction.
%
  Our formalisation is fully mechanised in Agda \cite{Zalakain2020Agda}.
\end{abstract}

@@

  \autoref{type-safety} details the type safety properties of our \picalc{} with leftovers.
  Leftover typing allows \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems to be stated.
  We use these to prove \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove \emph{subject reduction} (\autoref{thm:subject-reduction}).

@@

  \inferrule
  { }
  {\constr{scope-comm} : \new \new P \eqeq \new \new \func{swap}_{\constr{0}} \; P}
\end{mathpar}
\caption{Base cases of structural congruence.}
\label{fig:struct-cong-base}
\end{figure}

@@

In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof asserting that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$.
The function $\func{lower}_i \; Q \; uQ$ traverses $Q$ decrementing every index greater than $i$.
In $\constr{scope-comm}$ the function $\func{swap}_i \; P$ traverses $P$ (of type $\allowbreak \Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
In all the above, $i$ is incremented every time we go under a binder.

@@

In rule $\constr{comm}$, parallel processes reduce when they communicate over a common channel with index ${i}$.
As a result of that communication, the continuation of the input process $P$ has all the references to its most immediate variable substituted with references to $\suc j$, the variable sent by the output process $\send{i}{j}{Q}$.
After substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (we apply \autoref{lm:subst-unused} in \autoref{app:lemmas} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$).
Reduction is closed under parallel composition (rule $\constr{par}$), restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
--- notably, not under input nor output, as doing so would not preserve the sequencing of actions \cite{Sangio01}.

@@

\subparagraph*{Swapping}
of two variables preserves the well-typedness of a process.
We extend $\func{swap}_i$  introduced in \autoref{structural-congruence} to swap types in typing contexts and usage annotations in usage contexts.
\begin{nitheorem}[Swapping]
  \label{thm:swapping}
  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Then, $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\allowbreak \func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
\end{nitheorem}

@@

\begin{proof}[Proof (Sketch)]
  The proof is by induction on \textsc{Equals} $\eq{}$.
  For the nontrivial base cases and their symmetric variants, we apply a combination of \autoref{thm:framing}, \autoref{thm:strengthening}, \autoref{thm:weakening} and \autoref{thm:swapping}.
  Further details can be found in \autoref{app:struct} and the full proof in \cite{Zalakain2020Agda}.
\end{proof}

\subparagraph*{Substitution}
\autoref{thm:subst-generalization} proves a generalised version of substitution, where the substitution $\subst{P}{j}{i}$ is on any variable $i$.
\autoref{thm:substitution} instantiates the generalised version to the concrete case where $i$ is the most recently introduced variable $\constr{0}$, as required by subject reduction.

\begin{nitheorem}[Generalised substitution]
  \label{thm:subst-generalization}
  Let process $P$ be well typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
  Then, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
  \begin{multicols}{2}
  \begin{itemize}
    \item $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$
    \item $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$
  \end{itemize}
  \end{multicols}
  Let $\Gamma$ and $\Psi$ be such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
  Let $\Delta$ at position $i$ have usage $\lz$, meaning all consumption from $m$ to $n$ must happen in $P$.
  Then substituting $i$ with $j$ in $P$ will be well typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
\end{nitheorem}

\begin{proof}[Proof (Sketch)]
  By induction on $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
  Full details are found in \autoref{app:subst-generalization}.
  \begin{itemize}
   \item
     Constructor $\PO$: observe that $\Gamma_i \equiv \Psi_i$ and thus $\Gamma_j \equiv \Psi_j$.
   \item
     Constructor $\new$: we proceed inductively.
   \item
     Constructors $\recv{}{}$, $\send{}{}$ and $\comp{}{}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction. \qedhere
 \end{itemize}
\end{proof}

@@

      \draw[->] (gamma-i) -- node[above] {$\vdash P$} (xi-i);
      \draw[->] (xi-i) -- node[above] {$\vdash Q$} (psi-i);
      \draw[->,densely dotted] (gamma-j) -- node[below] {$\vdash \subst{P}{j}{i}$} (xi-j);
      \draw[->,densely dotted] (xi-j) -- node[below] {$\vdash \subst{Q}{j}{i}$} (psi-j);
      \draw[->,densely dotted] (xi-i) -- node[left] {$\ni_i l$} (theta);
      \draw[->,densely dotted] (xi-j) -- node[left] {$\ni_j l$} (theta);
    \end{tikzpicture}
    \caption{
      Diagrammatic representation of the $\comp{}{}$ case for substitution.
      Continuous lines represent known facts, dotted lines proof obligations.
    }
    \label{fig:subst}
\end{figure}

@@

      \draw[->] (gamma-i) -- (xi-j);
      \draw[->] (gamma-m) -- (xi-j);
      \draw[->,densely dotted] (gamma-j) -- (xi-j);
      \draw[->] (gamma-i) -- (gamma-m);
      \draw[->] (gamma-j) -- (gamma-m);
      \draw[->] (xi-i)    -- (xi-m);
      \draw[->] (xi-j)    -- (xi-m);
      \draw[->] (xi-i)    -- (psi-m);
      \draw[->] (xi-m)    -- (psi-m);
      \draw[->,densely dotted] (xi-j)    -- (psi-m);
    \end{tikzpicture}
    \caption{
      Alternative approach without arrows $\ni_i n$ and $\ni_j n$.
      This approach makes composing typing derivations before and after substitution more difficult.
    }
    \label{fig:subst-alternative}
\end{figure}

\begin{nitheorem}[Substitution]
  \label{thm:substitution}
  Let $P$ be well typed in $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
  Let $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
  Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
\end{nitheorem}
\begin{proof}[Proof (Sketch)]
  We use framing to derive $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$.
  Then, we use these to apply \autoref{thm:subst-generalization}.
\end{proof}

@@

\begin{proof}[Proof (Sketch)]
  By induction on $P \reduce{c} Q$.
  \hfill{}\\
  \begin{itemize}
    \item
    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) (to rearrange the assumptions), substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).

>>>>

  We present the syntax, semantics, type system and corresponding type safety properties.
  For the first time in the \picalc{}, we use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates linearity constraints into process continuations.
  We generalise the algebras on multiplicities, allowing the user to choose a mix of linear, graded and shared typing.
  We provide framing, weakening and strengthening theorems, which can now be stated for the linear \picalc{} and use them to prove subject congruence.
  We show that the type system is stable under renaming and prove subject reduction.
%
  Our formalisation is fully mechanised in Agda \cite{Zalakain2020Agda}.
\end{abstract}

@@

  \autoref{type-safety} details the type safety properties of our \picalc{} with leftovers.
  Leftover typing allows \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems to be stated.
  We use these to prove \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{renaming} (\autoref{thm:renaming}), to prove \emph{subject reduction} (\autoref{thm:subject-reduction}).

@@

  \inferrule
  { }
  {\constr{scope-comm} : \new \new P \eqeq \new \new \func{exchange}_{\constr{0}} \; P}
\end{mathpar}
\caption{Base cases of structural congruence.}
\label{fig:struct-cong-base}
\end{figure}

@@

In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof asserting that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$.
The function $\func{lower}_i \; Q \; uQ$ traverses $Q$ decrementing every index greater than $i$.
In $\constr{scope-comm}$ the function $\func{exchange}_i \; P$ traverses $P$ (of type $\allowbreak \Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
In all the above, $i$ is incremented every time we go under a binder.

@@

In rule $\constr{comm}$, parallel processes reduce when they communicate over a common channel with index ${i}$.
As a result of that communication, the continuation of the input process $P$ has all the references to its most immediate variable substituted with references to $\suc j$, the variable sent by the output process $\send{i}{j}{Q}$.
After this renaming, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (we apply \autoref{lm:renaming-unused} in \autoref{app:lemmas} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$).
Reduction is closed under parallel composition (rule $\constr{par}$), restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
--- notably, not under input nor output, as doing so would not preserve the sequencing of actions \cite{Sangio01}.

@@

\subparagraph*{Exchange}
of two variables preserves the well-typedness of a process.
We extend $\func{exchange}_i$  introduced in \autoref{structural-congruence} to exchange types in typing contexts and usage annotations in usage contexts.
\begin{nitheorem}[Exchange]
  \label{thm:exchange}
  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Then, $\types{\func{exchange}_i \; \gamma}{\func{exchange}_i \; \Gamma}{\allowbreak \func{exchange}_i \; P}{\func{exchange}_i \; \Xi}$.
\end{nitheorem}

@@

\begin{proof}[Proof (Sketch)]
  The proof is by induction on \textsc{Equals} $\eq{}$.
  For the nontrivial base cases and their symmetric variants, we apply a combination of \autoref{thm:framing}, \autoref{thm:strengthening}, \autoref{thm:weakening} and \autoref{thm:exchange}.
  Further details can be found in \autoref{app:struct} and the full proof in \cite{Zalakain2020Agda}.
\end{proof}

\subparagraph*{Renaming}
\autoref{thm:renaming-generalization} proves a generalised version of renaming, where the renaming $\subst{P}{j}{i}$ is on any variable $i$.
\autoref{thm:renaming} instantiates the generalised version to the concrete case where $i$ is the most recently introduced variable $\constr{0}$, as required by subject reduction.

\begin{nitheorem}[Generalised renaming]
  \label{thm:renaming-generalization}
  Let process $P$ be well typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
  Then, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
  \begin{multicols}{2}
  \begin{itemize}
    \item $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$
    \item $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$
  \end{itemize}
  \end{multicols}
  Let $\Gamma$ and $\Psi$ be such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
  Let $\Delta$ at position $i$ have usage $\lz$, meaning all consumption from $m$ to $n$ must happen in $P$.
  Then renaming $i$ to $j$ in $P$ will be well typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
\end{nitheorem}

\begin{proof}[Proof (Sketch)]
  By induction on $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
  Full details are found in \autoref{app:renaming-generalization}.
  \begin{itemize}
   \item
     Constructor $\PO$: observe that $\Gamma_i \equiv \Psi_i$ and thus $\Gamma_j \equiv \Psi_j$.
   \item
     Constructor $\new$: we proceed inductively.
   \item
     Constructors $\recv{}{}$, $\send{}{}$ and $\comp{}{}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction. \qedhere
 \end{itemize}
\end{proof}

@@

      \draw[->] (gamma-i) -- node[above] {$\vdash P$} (xi-i);
      \draw[->] (xi-i) -- node[above] {$\vdash Q$} (psi-i);
      \draw[->,densely dotted] (gamma-j) -- node[below] {$\vdash \subst{P}{j}{i}$} (xi-j);
      \draw[->,densely dotted] (xi-j) -- node[below] {$\vdash \subst{Q}{j}{i}$} (psi-j);
      \draw[->,densely dotted] (xi-i) -- node[left] {$\ni_i l$} (theta);
      \draw[->,densely dotted] (xi-j) -- node[left] {$\ni_j l$} (theta);
    \end{tikzpicture}
    \caption{
      Diagrammatic representation of the $\comp{}{}$ case for renaming.
      Continuous lines represent known facts, dotted lines proof obligations.
    }
    \label{fig:subst}
\end{figure}

@@

      \draw[->] (gamma-i) -- (xi-j);
      \draw[->] (gamma-m) -- (xi-j);
      \draw[->,densely dotted] (gamma-j) -- (xi-j);
      \draw[->] (gamma-i) -- (gamma-m);
      \draw[->] (gamma-j) -- (gamma-m);
      \draw[->] (xi-i)    -- (xi-m);
      \draw[->] (xi-j)    -- (xi-m);
      \draw[->] (xi-i)    -- (psi-m);
      \draw[->] (xi-m)    -- (psi-m);
      \draw[->,densely dotted] (xi-j)    -- (psi-m);
    \end{tikzpicture}
    \caption{
      Alternative approach without arrows $\ni_i n$ and $\ni_j n$.
      This approach makes composing typing derivations before and after renaming more difficult.
    }
    \label{fig:subst-alternative}
\end{figure}

\begin{nitheorem}[Renaming]
  \label{thm:renaming}
  Let $P$ be well typed in $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
  Let $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
  Then, we can rename the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
\end{nitheorem}
\begin{proof}[Proof (Sketch)]
  We use framing to derive $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$.
  Then, we use these to apply \autoref{thm:renaming-generalization}.
\end{proof}

@@

\begin{proof}[Proof (Sketch)]
  By induction on $P \reduce{c} Q$.
  \hfill{}\\
  \begin{itemize}
    \item
    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) (to rearrange the assumptions), renaming (\autoref{thm:renaming}) and strengthening (\autoref{thm:strengthening}).
